---
title: "p8105_hw2_xw2962"
author: "Xiaoyu Wu"
date: "2023-09-26"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(haven)
library(readxl)
```

## Problem One 

#### Step One: Clean Data in `pols-month.csv`

```{r,message=FALSE}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols_df = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv") |> 
  janitor::clean_names() |> 
# Read in dataset 
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
# seperate month into year, month_num and day. 
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
# create a variable called "president" by recoding prez_gop
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 
# arrange table and delete day and columns starts with "prez" 
pols_df
```

#### Step Two: Clean Data in `snp.csv`

```{r,message=FALSE}
snp_df = 
  read_csv("./fivethirtyeight_datasets/snp.csv") |> 
  janitor::clean_names() |> 
# Read in dataset 
  mutate(date=as.Date(date,format="%m/%d/%y"),
         date=as.Date(ifelse(date>Sys.Date(),
                             format(date,"19y-%m-%d"),
                             format(date)))) |>
         separate(date,into=c("year","month","day")) |> 
# Use separate() to break up the variable mon into integer variables year, month, and day
   mutate(month=month.abb[as.numeric(month)],
   year=as.numeric(year)) |>
# Replace month number with month name
  select(-day) |> 
# Remove the day variable
  arrange(year, month) |>
# Arrange according to year and month
  relocate(year, month) 
# Organize so that year and month are the leading columns

snp_df
```

#### Step Three: Tidy Data in `unemployment.csv`

```{r, message=FALSE}
unemployment_df = 
  read_csv("./fivethirtyeight_datasets/unemployment.csv") |> 
  janitor::clean_names() |>
# Read in dataset 
  pivot_longer(
   jan:dec,
   names_to = "month",
   values_to = "unemployment_num"
   ) 
# Use pivot_longer function to switch from “wide” to “long” format;
unemployment_df
```

#### Join Datasets

```{r}
pols_snp = 
  left_join(pols_df,snp_df,by=c("year","month")) |> 
  arrange(year,month)
# Join the datasets by merging snp into pols
pols_snp_unemployment=
  left_join(pols_snp,unemployment_df,by=c("year","month")) |> 
  arrange(year,month)
# Join unemployment_df to the result dataset from above
pols_snp_unemployment
```

#### Short Description of Dataset

Dataset "pols_df" has `r ncol(pols_df)` variables: `r colnames(pols_df)`. In this dataset, there are `r nrow(pols_df)` rows. And there are `r ncol(pols_df)` columns. Here we consider variables "gov_gop", "sen_gop", "rep_gop", "gov_dem", "sen_dem", "rep_dem" and "president" to be important for analysis.

Dataset "snp_df" has `r ncol(snp_df)` variables: `r colnames(snp_df)`. In this dataset, there are `r nrow(snp_df)` rows. And there are `r ncol(snp_df)` columns. Here we consider variables "close" to be important for analysis.

Dataset "unemployment_df" has `r ncol(unemployment_df)` variables: `r colnames(unemployment_df)`. In this dataset, there are `r nrow(unemployment_df)` rows. And there are `r ncol(unemployment_df)` columns. Here we consider variables "unemployment_num" to be important for analysis.

The joining dataset "pols_snp_unemployment" has `r ncol(pols_snp_unemployment)` variables: `r colnames(pols_snp_unemployment)`. In this dataset, there are `r nrow(pols_snp_unemployment)` rows. And there are `r ncol(pols_snp_unemployment)` columns. Here we consider variables "gov_gop", "sen_gop", "rep_gop", "gov_dem", "sen_dem", "rep_dem", "president","close" and "unemployment_num" to be important for analysis.

## Problem Two

#### Import, Clean, and Organize the Data for `Mr. Trash Wheel`
```{r}
mr_trash_wheel_df=
  read_excel("./Trash Wheel Collection Data.xlsx",range="A2:N586") |> 
# Specify the sheet in the Excel file and to omit non-data entries using arguments in read_excel
  janitor::clean_names() |>  
# Use reasonable variable names
  drop_na() |> 
# Omit rows that do not include dumpster-specific data
  mutate(
    homes_powered = ((weight_tons*500)/30
    )) |>
# Update the data to include a new homes_powered variable based on the noted calculation
  mutate(trash_wheel_name="mr_trash_wheel") |>
  relocate(trash_wheel_name)
# Add a variable for clear identification 
mr_trash_wheel_df
```

#### Import, Clean, and Organize the Data for `Professor Trash Wheel`

```{r}
professor_trash_wheel_df=
  read_excel("./Trash Wheel Collection Data.xlsx",2,range="A2:M108") |> 
# Specify the sheet in the Excel file and to omit non-data entries using arguments in read_excel
  janitor::clean_names() |>  
# Use reasonable variable names
  mutate(
    homes_powered = ((weight_tons*500)/30
    )) |> 
# Update the data to include a new homes_powered variable based on the noted calculation
  mutate(trash_wheel_name="professor_trash_wheel") |>
  relocate(trash_wheel_name) |> 
# Add a variable for clear identification
  mutate(year=as.character("year"))
professor_trash_wheel_df
```

#### Import, Clean, and Organize the Data for `Gwynnda Trash Wheel`

```{r}
Gwynnda_trash_wheel_df=
  read_excel("./Trash Wheel Collection Data.xlsx",4,range="A2:L157") |> 
# Specify the sheet in the Excel file and to omit non-data entries using arguments in read_excel
  janitor::clean_names() |>  
# Use reasonable variable names
  mutate(
    homes_powered = ((weight_tons*500)/30
    )) |> 
# Update the data to include a new homes_powered variable based on the noted calculation
  mutate(trash_wheel_name="Gwynnda_trash_wheel") |>
  mutate(year=as.character(year)) |>
  relocate(trash_wheel_name)
# Add a variable for clear identification
Gwynnda_trash_wheel_df
Gwynnda_trash_wheel_df_july=filter(Gwynnda_trash_wheel_df,month=="July" )
Gwynnda_trash_wheel_df_july_year=filter(Gwynnda_trash_wheel_df_july,year=="2021")
# Filter out the Gwynnda_trash_wheel data in july 
```

#### Combine these with the `Mr. Trash Wheel` Dataset to Create Single  Tidy Dataset

```{r}
mr_professor_Gwynnda=
  bind_rows(mr_trash_wheel_df,professor_trash_wheel_df,Gwynnda_trash_wheel_df) |>
  janitor::clean_names() |>
  select(trash_wheel_name,everything())

mr_professor_Gwynnda
```

#### Description
The data "mr_trash_wheel_df" is from the Mr.Trash Wheel sheet from May.2014 to July. 2022, which contains `r colnames(mr_trash_wheel_df)` and the number of observations is `r nrow(mr_trash_wheel_df)`. 

The data "professor_trash_wheel_df" is from the Professor Trash Wheel sheet from January.2017 to July. 2022, which contains `r colnames(professor_trash_wheel_df)` and the number of observations is `r nrow(professor_trash_wheel_df)`. 

The data "Gwynnda_trash_wheel_df" is from the Gwynnda Trash Wheel sheet from July.2021 to July. 2022, which contains `r colnames(Gwynnda_trash_wheel_df)` and the number of observations is `r nrow(Gwynnda_trash_wheel_df)`.

Dataset "mr_professor_Gwynnda" has 15 variables: `r colnames(mr_professor_Gwynnda)`. In this dataset, there are `r nrow(mr_professor_Gwynnda)` rows or observations. And there are `r ncol(mr_professor_Gwynnda)` columns or variables. 

Here, we consider "weight_tons","volume_cubic_yards" and "homes_powered" to be key variables. The variable "weight_tons" has mean `r mean(mr_professor_Gwynnda$weight_tons)` and standard deviation `r sd(mr_professor_Gwynnda$weight_tons)`. The variable "volume_cubic_yards" has mean `r mean(mr_professor_Gwynnda$volume_cubic_yards)` and standard deviation `r sd(mr_professor_Gwynnda$volume_cubic_yards)`. The variable "homes_powered" has mean `r mean(mr_professor_Gwynnda$homes_powered)` and standard deviation `r sd(mr_professor_Gwynnda$homes_powered)`. 

The total weight of trash collected by Professor Trash Wheel is `r sum(professor_trash_wheel_df$weight_tons)`. 

The total number of cigarette butts collected by Gwynnda in July of 2021 is `r sum(Gwynnda_trash_wheel_df_july_year$cigarette_butts)`.  

## Problem Three

#### Import, Clean and Tidy Dataset `MCI_baseline.csv`

```{r, warning=FALSE,message=FALSE}
mci_baseline_df = 
  read_csv("./data_mci/MCI_baseline.csv",skip=1) |> 
  janitor::clean_names() |>
# Read in dataset 
  mutate(sex = recode(sex, "1" = "male", "0" = "female")) |>
  mutate(apoe4 = recode(apoe4, "1" = "carrier", "0" = "non_carrier")) |>
# Recode sex and APOE4 carrier status
  mutate(age_at_onset=as.numeric(age_at_onset)) |>
  filter(age_at_onset!=".") |> 
# Remove any participants who do not meet the stated inclusion criteria
  rename("study_id" = "id") 

mci_baseline_df_noncleaned = 
  read_csv("./data_mci/MCI_baseline.csv",skip=1) |> 
  janitor::clean_names() |>
  mutate(sex = recode(sex, "1" = "male", "0" = "female")) |> 
  mutate(apoe4 = recode(apoe4, "1" = "carrier", "0" = "non_carrier"))  

mci_baseline_df_female=filter(mci_baseline_df_noncleaned,sex=="female" )
# Filter out participants who are females 
mci_baseline_df_female_carriers=filter(mci_baseline_df_female,apoe4=="carrier")
# Filter out participants who are females and who have apoe4 carriers 

mci_baseline_df
mci_baseline_df_noncleaned
```

#### Discussion 
We first imported and cleaned the dataset. Then we recoded the "sex" and "APOE4" into meaningful characters. And we removed any participants who do not meet the stated inclusion criteria. 

Dataset "mci_baseline_df" has `r ncol(mci_baseline_df)` variables: `r colnames(mci_baseline_df)`. 

In this dataset, there are `r nrow(mci_baseline_df_noncleaned)` participants. 
Of these, `r nrow(mci_baseline_df)` devlops MCI eventually. 

The average baseline age is `r mean(mci_baseline_df_noncleaned$current_age)`. 
`r ((nrow(mci_baseline_df_female_carriers))/(nrow(mci_baseline_df_female)))*100` % of women in the study are APOE4 carriers.    


#### Import, Clean and Tidy Dataset `mci_amyloid.csv`

```{r,message=FALSE}
mci_amyloid_df = 
  read_csv("./data_mci/mci_amyloid.csv",skip = 1) |> 
  janitor::clean_names() |>
# Read in dataset 
  drop_na() |>
# Remove any participants who do not meet the stated inclusion criteria
  pivot_longer(
   baseline:time_8,
   names_to = "time",
   values_to = "time_value"
   ) |>
# Use pivot_longer function to switch from “wide” to “long” format
   arrange(time,study_id)
# Arrange the dataset according to time and study_id

mci_amyloid_df_unpivot= 
  read_csv("./data_mci/mci_amyloid.csv",skip = 1) |> 
  janitor::clean_names() |>
# Read in dataset 
  drop_na()

mci_amyloid_df 
```


#### Discussion 
We first read in the dataset. And then we remove any participants who do not meet the stated inclusion criteria. Next, we use pivot_longer function to switch from “wide” to “long” format. Finally, we arrange the dataset according to time and study_id. 

Dataset "mci_amyloid_df" has `r ncol(mci_amyloid_df)` variables: `r colnames(mci_amyloid_df)`. In this dataset, there are `r nrow(mci_amyloid_df)` rows. 

## Join dataset 

```{r EVAL=FALSE}
baseline_amyloid= merge(mci_baseline_df,mci_amyloid_df_unpivot, 
             by="study_id")
# Calculate participants appear in both datasets 
unqiue_baseline=setdiff(mci_baseline_df_noncleaned$id, mci_amyloid_df_unpivot$study_id)
length(unqiue_baseline)
unqiue_amyloid=setdiff(mci_amyloid_df_unpivot$study_id, mci_baseline_df_noncleaned$id)
length(unqiue_amyloid)
# Calculatethe number of unqiue participants in each datasets. 
baseline_amyloid_merge= merge(mci_baseline_df,mci_amyloid_df, 
             by="study_id")
# Merge two datasets so that only participants who appear in both datasets are retained 
```

#### Discussion 

There are 146 unqiue partcipants in "MCI_baseline.csv" dataset and there are 10 unique participants in "mci_amyloid.csv" dataset. 

Only `r nrow(baseline_amyloid)` participants appear in both the baseline and amyloid datasets. 

For the resulting merged dataset, there are `r ncol(baseline_amyloid_merge)` variables, they are `r colnames(baseline_amyloid_merge)`.

#### Export Result as CSV

```{r}
write_csv(baseline_amyloid_merge,"baseline_amyloid_merge.csv")
```



